---
title: "PROJECT"
output: html_document
date: "2024-04-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
##Loading Libraries
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(reshape2)
```

##Loading the dataset
```{r}
biker<-read.csv(file.choose())
View(biker)

```

## Seeing the datastructure of the dataset
```{r}
str(biker)
```
#### We notice that there are 730 rows with 16 columns

## Checking NA for each columns
```{r}
colSums(is.na(biker))
```
#### We can see there are no NA values


## Changing the datatype of dteday into date format
```{r}
biker$dteday <- as.Date(biker$dteday, format="%d-%m-%Y")
```


```{r}
str(biker)
```
#### We can see that the format for each column is correct and can be directly utlized in data processing

## Removing yr column
```{r}
biker[, 'yr']<- NULL
```

```{r}
head(biker)
```
#### The column biker has been dropped

## Seeing the outliers in each column 
```{r}
x <- melt(biker)

plt <- ggplot(data = x, aes(x = variable, y = value))
plt + geom_boxplot() + theme_minimal() + labs(x = "Title", y = "x")
```
#### We can see that clearly there are outliers

## Analysing one by one
```{r}
lapply(seq_along(biker), function(x){
 boxplot(biker[[x]], main = paste("Title", colnames(biker))[[x]])
})
```

#### We can see that all values in variables are in with in expected values. Moving forward to Exploratory Data Analysis

Linear Regression Analysis

Creating a linear regression model for 'registered' variable
```{r}
registered_model <- lm(registered ~ ., data = biker)
summary(registered_model)

```

Creating a linear regression model for 'casual' variable
```{r}
casual_model <- lm(casual ~ ., data = biker)
summary(casual_model)

```

Creating a linear regression model for 'count' variable
```{r}
count_model <- lm(cnt ~ ., data = biker)
summary(count_model)
```

Performing Time-Series Analysis to plot historical trends of bike share users over time
```{r}
ggplot(data = biker, aes(x = dteday, y = cnt)) +
  geom_line() +
  labs(x = "Date", y = "Count of Bike Share Users") +
  ggtitle("Historical Trends of Bike Share Users Over Time")
```
The above plot clearly shows that there is an overall upward trend in the total number of bike share users between 2018 and 2020.

Explore distribution of bike share users across various factors

Seasons
```{r}
ggplot(data = biker, aes(x = factor(season), y = cnt)) +
  geom_boxplot() +
  labs(x = "Season", y = "Count of Bike Share Users") +
  ggtitle("Distribution of Bike Share Users by Season")
```

Months
```{r}
ggplot(data = biker, aes(x = factor(mnth), y = cnt)) +
  geom_boxplot() +
  labs(x = "Month", y = "Count of Bike Share Users") +
  ggtitle("Distribution of Bike Share Users by Month")

```

Weekdays
```{r}
ggplot(data = biker, aes(x = factor(weekday), y = cnt)) +
  geom_boxplot() +
  labs(x = "Weekday", y = "Count of Bike Share Users") +
  ggtitle("Distribution of Bike Share Users by Weekday")

```

Holidays
```{r}
ggplot(data = biker, aes(x = factor(holiday), y = cnt)) +
  geom_boxplot() +
  labs(x = "Holiday", y = "Count of Bike Share Users") +
  ggtitle("Distribution of Bike Share Users on Holidays")
```



```{r}
# Scatter plot with regression lines for temperature and bike share usage
ggplot(data = biker, aes(x = temp, y = cnt)) +
  geom_point(alpha = 0.5) +  # Add points
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  
  labs(x = "Temperature", y = "Count of Bike Share Users") +
  ggtitle("Relationship Between Temperature and Bike Share Usage")

# Scatter plot with regression lines for humidity and bike share usage
ggplot(data = biker, aes(x = hum, y = cnt)) +
  geom_point(alpha = 0.5) +  # Add points
  geom_smooth(method = "lm", se = FALSE, color = "green") +
  labs(x = "Humidity", y = "Count of Bike Share Users") +
  ggtitle("Relationship Between Humidity and Bike Share Usage")

# Scatter plot with regression lines for windspeed and bike share usage
ggplot(data = biker, aes(x = windspeed, y = cnt)) +
  geom_point(alpha = 0.5) +  # Add points
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(x = "Windspeed", y = "Count of Bike Share Users") +
  ggtitle("Relationship Between Windspeed and Bike Share Usage")

```

